---
title: "DATA 607 - Assignment 7 - Web Technologies; MongoDB"
author: "Preston Peck"
date: "10/10/2021"
output: html_document
---

```{r include=FALSE}
library(dplyr)
library(jsonlite)
library(rvest)
library(readr)
library(stringr)
library(xml2)
library(XML)
library(tidyr)
```

# Load JSON
```{r}
# Variables
jsonFile <- "data/books.json"

# Convert to Data Frame
jsonTable <- fromJSON(jsonFile) %>%
  as_tibble()

# Tidy
jsonTable$link <- jsonTable$link %>%
  str_trim()

jsonTable
```

# Load XML
### XML Library
```{r}
# Variables
nodeName <- "//book"
xmlFile <- "data/books.xml"

# Parse Data
xmlData <- xmlFile %>%
  xmlParse()

nodeData <- xmlData %>%
  getNodeSet(nodeName)

# Convert to Data Frame
xmlTable <- xmlData %>%
  xmlToDataFrame(nodes = nodeData) %>%
  as_tibble()

# Tidy
xmlTable$link <- xmlTable$link %>%
  str_trim()

xmlTable$pages <- xmlTable$pages %>%
  as.integer()

xmlTable$year <- xmlTable$year %>%
  as.integer()

xmlTable
```
### xml2 Library
```{r}
# Parse Data
xml2Data <- xmlFile %>%
  read_xml()
```
##### All Nodes
```{r}
# Parse Data
nodeData <- xml2Data %>%
  xml_find_all(nodeName)
nodeData

nodeLengths <- nodeData %>%
  xml_length()

nodeNumbers <- 1:length(nodeData)

nodeNames <- nodeData %>%
  xml_children() %>%
  xml_name()

nodeValues <- nodeData %>%
  xml_children() %>%
  xml_text()

# Convert to Data Frame
xml2AllTable <- tibble(
  num = rep(nodeNumbers, times = nodeLengths),
  name = nodeNames, 
  value = nodeValues
)

xml2AllTable

# Tidy
xml2AllTable <- (xml2AllTable %>%
  spread(name, value))[,-c(1)]
  
xml2AllTable$link <- xml2AllTable$link %>%
  str_trim()

xml2AllTable$pages <- xml2AllTable$pages %>%
  as.integer()

xml2AllTable$year <- xml2AllTable$year %>%
  as.integer()

xml2AllTable
```
##### Select Nodes
```{r}
# Variables
authorName <- "author"
countryName <- "country"
imageLinkName <- "imageLink"
languageName <- "language"
linkName <- "link"
pagesName <- "pages"
titleName <- "title"
yearName <- "year"

# Parse Data
authorNodes <- nodeData %>%
  xml_find_all(xpath = authorName) %>%
  xml_text()

countryNodes <- nodeData %>%
  xml_find_all(xpath = countryName) %>%
  xml_text()

imageLinkNodes <- nodeData %>%
  xml_find_all(xpath = imageLinkName) %>%
  xml_text()

languageNodes <- nodeData %>%
  xml_find_all(xpath = languageName) %>%
  xml_text()

linkNodes <- nodeData %>%
  xml_find_all(xpath = linkName) %>%
  xml_text()

pagesNodes <- nodeData %>%
  xml_find_all(xpath = pagesName) %>%
  xml_integer()

titleNodes <- nodeData %>%
  xml_find_all(xpath = titleName) %>%
  xml_text()

yearNodes <- nodeData %>%
  xml_find_all(xpath = yearName) %>%
  xml_integer()

# Convert to Data Frame
xml2SelectTable <- tibble(
  author = authorNodes,
  country = countryNodes,
  imageLink = imageLinkNodes,
  language = languageNodes,
  link = linkNodes,
  pages = pagesNodes,
  title = titleNodes,
  year = yearNodes
)

# Tidy
xml2SelectTable$link <- xml2SelectTable$link %>%
  str_trim()

xml2SelectTable
```

# Load HTML Table
```{r}
# Variables
htmlFile <- "data/books.html"
tableName <- "table"

# Convert to Data Frame
htmlTable <- read_html(htmlFile) %>%
  html_element(tableName) %>%
  html_table()

htmlTable
```

# Compare Results
```{r}
identical(jsonTable, xmlTable)
identical(xmlTable, xml2AllTable)
identical(xml2AllTable, xml2SelectTable)
identical(xml2SelectTable, htmlTable)
```

# Analysis
All three formats can eventually be formatted into the same data frame, but without going outside of each respective formats' library, the link column ended up with a new line character at in the end in JSON, and a whitespace character at the end in XML. The XML library also had difficulty determining data type and treated everything as text. As a result, HTML appears to be the easiest to work with, followed by JSON, and then XML.